 

### 2019-07-26 Manuscript/SupplementaryMaterial.tex
Update supplementary material

In this Supplementary Material we give a more detailed description of
the cylindrical regression models and outline the MCMC procedures to fit
them. R-code for the MCMC sampler and the analysis of the teacher data
can be found here:
\url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.
Note that the dimensions of the objects (design matrices, mean vectors,
etc.) are those that were used in the analysis of the teacher data where
we have 1 circular outcome, 1 linear outcome and estimate an intercept
and regression coefficient for the covariate self-efficacy. Note that
for the regression of the linear component in the CL-PN and CL-GPN
models we also have the sine and cosine of the circular outcome in the
regression equation, this makes the vector with regression coefficients,
\(\boldsymbol{\gamma}\), four-dimensional.

@@

Following Mastrantonio, Maruotti, \& Jona-Lasinio (2015) we consider in
this section two models where the relation between
\(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\) and \(q\)
covariates is specified as \begin{equation}\label{circlinlink}
Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
\end{equation} \noindent where the random variable \(R\geq0\) will be
introduced below, the error term \(\epsilon \sim N(0, \sigma^2)\) with
variance \(\sigma^2>0\),
\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
the intercept and regression coefficients and \(x_1, \dots, x_q\) are
the \(q\) covariates. In both of these models the conditional
distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given by
\begin{equation}\label{ycondtheta}
f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
\end{equation} \noindent where
\(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta) \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
\(r \geq 0\). The linear outcome thus has a normal distribution
conditional on \(\Theta\) and \(R\) and contains already linear
covariates \(x_1, \dots, x_q\) in its location part.\newline \indent For
the circular outcome we assume either a projected normal (PN) or a
general projected normal (GPN) distribution. These distributions arise
from the radial projection of a distribution defined on the plane onto
the circle. The relation between a bivariate vector \(\boldsymbol{S}\)
in the plane and the circular outcome \(\Theta\) is defined as follows
\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
\end{equation} \noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\),
the Euclidean norm of the bivariate vector \(\boldsymbol{S}\). In the PN
distribution we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the
GPN we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
\(\boldsymbol{\mu} \in \mathbb{R}^2\),
\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\),
and \(\xi,\tau \in (-\infty, +\infty)\) (as in Hernandez-Stumpfhauser,
Breidt, \& \VANDER{Woerd}{Van der}{van der} Woerd (2016)). This leads to
the circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN)
distributions. We will now detail how we modify both cylindrical
distributions to also incorporate covariates for the circular part.

@@

Following Nuñez-Antonio, Gutiérrez-Peña, \& Escarela (2011), the joint
density of \(\Theta\) and \(R\) for the PN distribution equals
\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\mu^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(u^t\mu) \right]\},
\end{equation} \noindent where
\(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta) \end{bmatrix}\)
and \(r\) is the same as in \eqref{circlinlink} and \eqref{ycondtheta}
and is defined in \eqref{projection}. In a regression setup the outcomes
\(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where \(n\) is
the sample size, are generated independently from the distribution with
density \eqref{pnreg}. The mean vector
\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is then defined as
\(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\) where the
vector \(\boldsymbol{z}_i\) is a vector of dimension \(p + 1\) that
contains the covariate values and the value 1 to estimate an intercept
and
\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
contains the regression coefficients and intercepts.

@@

Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
(2016) the joint density of \(R\) and \(\Theta\) for the GPN
distribution equals \begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2\tau^2}\right],
\end{equation} \noindent where we recall that
\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\).
In a regression setup the outcomes \(\theta_i\) and \(r_i\) for each
individual are generated independently from \eqref{gpnreg}. The mean
vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined in the same
way via covariates as for the modified CL-PN distribution.

@@

Both cylindrical models introduced here are estimated using Markov Chain
Monte Carlo (MCMC) methods based on Nuñez-Antonio et al. (2011), Wang \&
Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
regression of the circular outcome.

@@

This model is an extension of the cylindrical model introduced in Abe \&
Ley (2017) to the regression context. The joint density of \(\Theta\)
and \(Y\), in this model defined only on the positive real half-line
\([0, + \infty)\), reads \begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation} \noindent where \(\alpha > 0\) is a linear shape
parameter, \(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular
concentration and skewness parameters with \(\kappa\) also regulating
the circular-linear dependence. Our modification occurs at the level of
the linear scale parameter \(\nu>0\) and circular location parameter
\(\mu\in [0, 2\pi)\), both of which we express in terms of covariates:
\(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
coefficients \(\gamma_j \in (-\infty, +\infty)\) for the prediction of
\(y\) where \(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The
parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and
\(\boldsymbol{\beta}\) is a vector of \(p\) regression coefficients
\(\beta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\)
where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector
of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each
individual is generated independently from the modified density
\eqref{WeiSSVMdensity}.\newline \indent As in Abe \& Ley (2017), the
conditional distribution of \(Y\) given \(\Theta=\theta\) is a Weibull
distribution with shape \(\alpha\) and scale
\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the
conditional distribution of \(\Theta\) given \(Y=y\) is a sine skewed
von Mises distribution with location parameter \(\mu\) and concentration
parameter \((\nu y)^\alpha\tanh(\kappa)\). The log-likelihood for this
model equals \begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))).\nonumber
\end{align} \noindent We can use numerical optimization (Nelder-Mead) to
find solutions for the maximum likelihood (ML) estimates for the
parameters of the model.

@@

independent circular outcomes and \(w\) independent linear outcomes. The
circular outcomes

@@

circular outcomes. The linear outcomes

@@

the linear outcomes are represented as

@@

outcome \(\boldsymbol{Y}\),

@@

outcome is created by modelling the augmented representations of

@@

the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)

@@

circular outcome and the linear outcome. \newline \indent In our

@@

circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean
in the density in \eqref{YDThetarjoint} then becomes

@@

$$l(\theta, r) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\mu}_i\Sigma^{-1}(u_i^t\hat{\mu}_i)^t)/2\tau^2$$
where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\mu}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.

@@

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}_{n \times 2}$ in a training (90\%) and holdout (10\%) set. Note that in this paper we have only one circular outcome and one linear outcome and the MCMC procedure outlined here is specified for this situation. It can however be generalized to a situation with multiple circular and linear outcomes without too much effort. 
\item Define the prior parameters for the training set. Since we have only one circular outcome, one linear outcome and one covariate, we have $m = 1$, $w = 1$ and $g = 1$. In this paper we use the following priors:

@@

\item Set starting values $\boldsymbol{\beta} = (0,0,0,0,0,0)^t$, $\boldsymbol{\Sigma} = \boldsymbol{I}_3$ and $\lambda = 0$. Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Compute the latent outcomes $\tilde{y}_i$ underlying the linear outcome for the holdout and training dataset as follows:
$$\tilde{y}_i = \lambda d_i. $$
\item Compute $\boldsymbol{\eta}_i$ defined as follows for each individual $i$:
$$\boldsymbol{\eta}_i = (\boldsymbol{s}_i^t,y_i)^t - (\boldsymbol{0}_{2m}^t, \lambda d_i)^{t}.$$

@@

\DeclareRobustCommand{\VANDER}[3]{#3}
\DeclareRobustCommand{\VAN}[3]{#3}
\DeclareRobustCommand{\DEN}[3]{#3}

>>>>

In this Supplementary Material we give a more detailed description of
the cylindrical regression models and outline the MCMC procedures to fit
them. R-code for the MCMC sampler and the analysis of the teacher data
can be found here:
\url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.
Note that the dimensions of the objects (design matrices, mean vectors,
etc.) are those that were used in the analysis of the teacher data where
we have 1 circular component, 1 linear component and estimate an
intercept and regression coefficient for the covariate self-efficacy.
Note that for the regression of the linear component in the CL-PN and
CL-GPN models we also have the sine and cosine of the circular component
in the regression equation, this makes the vector with regression
coefficients, \(\boldsymbol{\gamma}\), four-dimensional.

@@

Following Mastrantonio, Maruotti, \& Jona-Lasinio (2015) we consider in
this section two models where the prediction equation for the linear
component is specified as \begin{equation}\label{linpredCLPNCLGPN}
\hat{y_i} = \gamma_0 + \gamma_{cos}*\cos(\theta_i)*r_i + \gamma_{sin}*\sin(\theta_i)*r_i + \gamma_1*x_1 + \dots + \gamma_q*x_q,
\end{equation} \noindent where \(r_i\) is a realization of the
unobserved the random variable \(R\geq0\) that will be introduced below,
\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
the intercept and regression coefficients and \(x_1, \dots, x_q\) are
the \(q\) covariates. In both of these models the conditional
distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given by
\begin{equation}\label{ycondtheta}
f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
\end{equation} \noindent where
\(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta) \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
\(r \geq 0\). The linear component thus has a normal distribution
conditional on \(\Theta\) and \(R\) and contains already linear
covariates \(x_1, \dots, x_q\) in its location part.\newline \indent For
the circular component we assume either a projected normal (PN) or a
general projected normal (GPN) distribution. These distributions arise
from the radial projection of a distribution defined on the plane onto
the circle. The relation between a bivariate vector \(\boldsymbol{S}\)
in the plane and the circular component \(\Theta\) is defined as follows
\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
\end{equation} \noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\),
the Euclidean norm of the bivariate vector \(\boldsymbol{S}\). In the PN
distribution we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the
GPN we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
\(\boldsymbol{\mu} \in \mathbb{R}^2\),
\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\),
and \(\xi,\tau \in (-\infty, +\infty)\) (as in Hernandez-Stumpfhauser,
Breidt, \& \VANDER{Woerd}{Van der}{van der} Woerd (2016)). This leads to
the circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN)
distributions. We will now detail how we modify both cylindrical
distributions to also incorporate covariates for the circular part.

@@

Following Nuñez-Antonio, Gutiérrez-Peña, \& Escarela (2011), the joint
density of \(\Theta\) and \(R\) for the PN distribution equals
\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{1}{2\pi} \exp\{-0.5 \mid\mid\boldsymbol{\mu}^2\mid\mid\}\exp\{-0.5\left[r^2 -2r(\boldsymbol{u}^t\boldsymbol{\mu}) \right]\},
\end{equation} \noindent where
\(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta) \end{bmatrix}\)
and \(r\) is defined in \eqref{projection}. In a regression setup the
outcome components \(\theta_i,r_i\) for each individual
\(i = 1, \dots, n\), where \(n\) is the sample size, are generated
independently from the distribution with density \eqref{pnreg}. The mean
vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is then defined as
\(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\) where the
vector \(\boldsymbol{z}_i\) is a vector of dimension \(p + 1\) that
contains the covariate values and the value 1 to estimate an intercept
and
\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
contains the regression coefficients and intercepts.

@@

Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
(2016) the joint density of \(R\) and \(\Theta\) for the GPN
distribution equals \begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2\tau^2}\right],
\end{equation} \noindent where we recall that
\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}\).
In a regression setup the outcome components \(\theta_i\) and \(r_i\)
for each individual are generated independently from \eqref{gpnreg}. The
mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined in the
same way via covariates as for the modified CL-PN distribution.

@@

Both cylindrical models introduced here are estimated using Markov Chain
Monte Carlo (MCMC) methods based on Nuñez-Antonio et al. (2011), Wang \&
Gelfand (2013) and Hernandez-Stumpfhauser et al. (2016) for the
regression of the circular component.

@@

This model is an extension of the cylindrical model introduced in Abe \&
Ley (2017) to the regression context. The joint density of \(\Theta\)
and \(Y\), in this model defined only on the positive real half-line
\([0, + \infty)\), reads \begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu_c))
                 y^{\alpha-1}
                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu_c))],
\end{equation} \noindent where \(\alpha > 0\) is a linear shape
parameter, \(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular
concentration and skewness parameters with \(\kappa\) also regulating
the circular-linear dependence. Our modification occurs at the level of
the linear scale parameter \(\nu>0\) and circular location parameter
\(\mu_c\in [0, 2\pi)\), both of which we express in terms of covariates:
\(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\mu_{c,i} = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\).
The parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression
coefficients \(\gamma_j \in (-\infty, +\infty)\) for the prediction of
\(y\) where \(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The
parameter \(\beta_0 \in [0, 2\pi)\) is the intercept and
\(\boldsymbol{\beta}\) is a vector of \(p\) regression coefficients
\(\beta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\)
where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector
of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome component vector \((\theta_i, y_i)^t\) for
each individual is generated independently from the modified density
\eqref{WeiSSVMdensity}.\newline \indent As in Abe \& Ley (2017), the
conditional distribution of \(Y\) given \(\Theta=\theta\) is a Weibull
distribution with shape \(\alpha\) and scale
\(\nu(1-\tanh(\kappa)\cos(\theta - \mu_c))^{1/\alpha}\) and the
conditional distribution of \(\Theta\) given \(Y=y\) is a sine skewed
von Mises distribution with location parameter \(\mu_c\) and
concentration parameter \((\nu y)^\alpha\tanh(\kappa)\). The
log-likelihood for this model equals
\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - (\beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})))).\nonumber
\end{align} \noindent We can use numerical optimization (Nelder-Mead) to
find solutions for the maximum likelihood (ML) estimates for the
parameters of the model.

@@

independent circular components and \(w\) independent linear components.
The circular components

@@

circular components. The linear components

@@

the linear components are represented as

@@

component \(\boldsymbol{Y}\),

@@

component is created by modelling the augmented representations of

@@

the circular component and the matrix \(\boldsymbol{\Sigma}_{sy}\)

@@

circular component and the linear component. \newline \indent In our

@@

circular components, \(w\) linear components and \(g\) covariates. The
mean in the density in \eqref{YDThetarjoint} then becomes

@@

$$l(\theta, r) =  \log(1) - \log(2\pi+\tau) - \sum \log(r_i) + (u_i^t\hat{\boldsymbol{\mu}}_i\Sigma^{-1}(u_i^t\hat{\boldsymbol{\mu}}_i)^t)/2\tau^2$$
where $u_i = (cos\theta_i, \sin \theta_i)$ and $\hat{\boldsymbol{\mu}}_i = (\beta_0^{I} + \beta_0^{I}\text{SE}_i, \beta_0^{II} + \beta_0^{II}\text{SE}_i)^t$.

@@

\begin{enumerate}
\item Split the data, with the circular component $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear component $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}_{n \times 2}$ in a training (90\%) and holdout (10\%) set. Note that in this paper we have only one circular component and one linear component and the MCMC procedure outlined here is specified for this situation. It can however be generalized to a situation with multiple circular and linear components without too much effort. 
\item Define the prior parameters for the training set. Since we have only one circular component, one linear component and one covariate, we have $m = 1$, $w = 1$ and $g = 1$. In this paper we use the following priors:

@@

\item Set starting values $\boldsymbol{\beta} = (0,0,0,0,0,0)^t$, $\boldsymbol{\Sigma} = \boldsymbol{I}_3$ and $\lambda = 0$. Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate scores $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular component for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Compute the latent scores $\tilde{y}_i$ underlying the linear component for the holdout and training dataset as follows:
$$\tilde{y}_i = \lambda d_i. $$
\item Compute $\boldsymbol{\eta}_i$ defined as follows for each individual $i$:
$$\boldsymbol{\eta}_i = (\boldsymbol{s}_i^t,y_i)^t - (\boldsymbol{0}_{2m}^t, \lambda d_i)^{t}.$$

@@
